Who introduced this bug? \\ It may not have been caused by the previous commit!

### 2016-04-10 paperICSME.tex
Intro reviewed.

This paper presents an observational study designed to sheed some more light in this area. For that, we studied the lines changed by bug fixes as a part of ``the previous commit'' (or commits) in the OpenStack project. Using information from the code management, issue tracking, and code review systems, we analyzed if the code introduced by previous commits was correct at the time of introduction, and was therefore the cause of the bug. Our results show that the assumption that bugs were introduced in the previous commit does not hold for a large fraction (at least 37\%) of the analyzed bugs.
\end{abstract}

@@

When a failure is found in some software, developers try to fix it by locating and modifying the source code line(s) that are the cause for the wrong behavior. It may seem at first reasonable to assume that previous modifications of this line or lines are the cause of the bug. That previous modification is what will refer through this paper as the \textit{previous commit}.

But in fact, to find when and where a bug was introduced in the source code is not a trivial task, and by far more complex than this assumption. This has been largely ignored in much of the bug-fix literature, mainly because the data related to the origin of a bug is \emph{embedded} in the evolution of the software~\cite{sinha2010buginnings}, as Sinha et al. state. With this they mean that there is no easy evidence (artifact, comment or log) where developers specify what produced that error from a more historical point of view. The explanation of the cause is thus \emph{embedded} in the project.

This is the reason why many studies in the area of software maintenance and evolution start with this implicit assumption. As an anecdotal evidence, we have found the following rationales in several areas of research:

@@

But although the assumption can be found frequently in the research literature, in our opinion there is not enough empirical evidence supporting it. That is the reason why we have conducted an observational study on bug fixing, devoting a significant effort to locate the origin of a bug in the source code and understanding the possible causes. For this we took into account when the line was inserted and the general context of the project at that point.

Figure~\ref{fig:1} shows a clear example of what we understand as the cause of the bug. Let's assume that we have three different versions of the same file in the history of the control version of the project.

@@

\begin{enumerate}
  \item The code on the left (subfigure \textit{(1)}) is the one written to fix the bug.
  \item The code in the middle (subfigure \textit{(2)}) shows the moment in which the bug was introduced (being \textit{31f08423} the id of change), the previous commit.
  \item The code on the right (subfigure \textit{(3)}) shows how in previous versions of the file, the bug did not exist.
\end{enumerate}

@@

According to the description in the log of the commit that fixed the bug (see Figure~\ref{fig:2}), commit \textit{31f08423} was the one where the bug was introduced, as it used a string variable, while a Boolean had to be used, keeping the concordance with the rest of the code. Therefore, in this case, the bug was introduce in the previous commit.

On the other hand, Figure~\ref{fig:3} shows a clear example of a case where the cause of the bug cannot be attributed to the previous commit. In this example, the bug fixing commit log (see Figure~\ref{fig:4}) describes that the name of an argument changed when updating the version causing the failure in the software. This change was done because of new requirements in the software version, and is unrelated to the modifications in the previous commit. When the modified lines where introduced the first time, they where not buggy.

@@

Based on anecdotal evidence like the one presented in Figures~\ref{fig:3} and~\ref{fig:4}, we argue that in projects that are continuously evolving, with a large development community, code that at some point was correct could become buggy later. Changes in other parts
of the code may trigger wrong behavior (bugs) in places which were correct in the past. This happens often in situations like changes of the API. In the moment the code was written, it was correct and the software worked fine. Additions of new features or enhancements to the API may have as a side effect that the formerly correct code starts to show a wrong behavior, making the software fail. In such cases, the source of the error cannot be attributed to the changes performed in the previous commit, which were correct when they were introduced, since in that moment they referred to a different API.

@@

In detail, we attempt to address the following research questions:

\begin{itemize}
     \item RQ1: How can we identify changes done to fix a bug?
     \item RQ2: How often is the previous commit the cause of the bug?
\end{itemize}

RQ2 is the main question that we want to answer in this paper: given bugs that have been fixed by changing some code, how many of those were introduced by the previous commit?

But be able to answer RQ2, we first need to study the issue-tracking system and identify the subset of closed tickets that correspond to fixed bugs. In essence, RQ1 could be also stated as \emph{Which tickets in the issue tracking system are (real) bug reports?}. This is because (real) bugs are managed in an issue-tracking system together with feature requests, optimization, test cases, etc. As we are only interested in bugs, we need to first identify those as a previous step to analyze if they have been caused by the previous commit.

One interesting aspect of our study is that it addresses a very fundamental aspect of many studies on how bugs are fixed: the underlying assumption that for fixed code there must be a commit previous to the fix that \emph{touched} the same lines (in a wrong way). If evidence is found that in a large fraction of the cases the corresponding code was correct when it was introduced, there is no reason to blame it as the cause of the bug, even when changing it fixes the bug. Therefore, any result obtained after this assumption should be revisited with some care. We want to contribute with a first step in removing this uncertainty.

>>>>

This paper presents an observational study designed to shed some more light in this area. For that, we studied the lines changed by bug fixes as a part of ``the previous commit'' (or commits) in the OpenStack project. Using information from the code management, issue tracking, and code review systems, we analyzed if the code introduced by previous commits was correct at the time of introduction, and was therefore the cause of the bug. Our results show that the assumption that bugs were introduced in the previous commit does not hold for a large fraction (at least 37\%) of the analyzed bugs.
\end{abstract}

@@

When a failure is found in some software, developers try to fix it by locating and modifying the source code line(s) that are the cause for the wrong behavior. It may seem reasonable to assume that previous modifications of these lines are the cause of the bug. Those previous modifications are what will refer through this paper as the \textit{previous commits}.

However, to find where and when a bug was introduced in the source code is not a trivial task, and it may be much more complex than what this assumption suggests. This fact has been largely ignored in much of the bug-fix literature, mainly because the data related to the origin of a bug is \emph{embedded} in the evolution of the software~\cite{sinha2010buginnings}, as Sinha et al. state. With this they mean that there is no easy evidence (artifact, comment or log) where developers specify what produced that error from a more historical point of view. The explanation of the cause is thus \emph{embedded} in the project.

This is the reason why many studies in the area of software maintenance and evolution start with this implicit assumption. As an anecdotal evidence, we have found the following rationales in several research areas:

@@

But although the assumption can be found frequently in the research literature, in our opinion there is not enough empirical evidence supporting it. That is the reason why we have conducted an observational study on bug fixing, devoting a significant effort to locate the origin of bugs in source code, and to understand the possible causes. Our main aim was to identify it the previous commit to to a bug fix was buggy ``at the time of committing it''. That is, if the program showed the bug at that moment because of the code introduced in that commit, or on the contrary, if the bug was caused by some earlier commit to that code, or by some other commit affecting other parts of the program.

To illustrate a case in which the assumption of the previous commit is fulfilled, Figure~\ref{fig:1} shows an example of what we understand as a previous commit causing a bug. Let's assume that we have three different versions of the same file in the history of the control version of the project:

@@

\begin{enumerate}
  \item The code on the left (subfigure \textit{(1)}) is the one after the bug was fixed.
  \item The code in the middle (subfigure \textit{(2)}) shows the previous version of the code (the previous commit), which was already causing malfunction.
  \item The code on the right (subfigure \textit{(3)}) shows the version of the file before the previous commit. In this case, at that point the bug did not exist.
\end{enumerate}

@@

According to the description in the log of the commit that fixed the bug (see Figure~\ref{fig:2}), the previous commit introduced it: a string variable was produced instead of a Boolean variable, which was needed to satisfy expectations by the rest of the code.

For an example of the contrary, when the assumption is not fulfilled, consider Figure~\ref{fig:3}, which shows a case in which the code introduced by the previous commit was correct at the moment of introducing it. Now, the log of the commit fixing the bug (Figure~\ref{fig:4}) describes that the name of an argument had changed due to some other commit to other part of the software, causing the failure. In this case, the failure has nothing to do with the previous commit, but with later changes in other areas that were not properly reflected in this code. When the lines were introduced by the previous commit, they where not buggy.

@@

This is an example of how in projects that are evolving, code that at some point was correct could become buggy later. Changes in other parts of the code may trigger bugs in places which were correct in the past. This happens often in situations of changes in called APIs: when the code was written, it was correct, although later modifications to APIs had as a side effect that the formerly correct code started to show a wrong behavior, making the software fail. In such cases, the source of the error cannot be attributed to the changes performed in the previous commit, which were correct when they were introduced, since in that moment they referred to a different API.

@@

More formally, we address the following research questions:

\begin{itemize}
     \item RQ1: How can we identify when changes to code do fix a bug?
     \item RQ2: How often is the previous commit the cause of a bug?
\end{itemize}

RQ2 is the main question that we want to answer: given bugs that have been fixed by changing some code, how many of those were introduced by the previous commit?

In practice, to answer RQ2 we first need to study the issue-tracking system and identify the subset of closed tickets that correspond to fixed bugs. In essence, RQ1 could be also stated as \emph{Which tickets in the issue tracking system are (real) bug reports?}, assuming that we can track tickets to changes in code. This is because (real) bugs are managed in an issue-tracking system together with feature requests, optimization, test cases, etc. As we are only interested in bugs, we need to first identify those as a previous step to analyze if they have been caused by the previous commit.

While RQ1 is instrumental, with RQ2 we address a very fundamental aspect of the studies on how bugs are fixed: the underlying assumption that the commit previous to a bug fix \emph{touched} the same lines introducing the bug. If evidence is found that in a large fraction of the cases the corresponding code was correct when it was introduced, there is no reason to blame it as the cause of the bug, even if changing it fixes the bug. Therefore, any result obtained after this assumption should be revisited with some care.
